### 2025-02-13 brms_meeg.tex
pushing some preliminary simulation results

\documentclass[
  doc,
  floatsintext,
  longtable,
  a4paper,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\geometry{inner=1in, outer=1in}
\fancyhfoffset[LE,RO]{0cm}

\usepackage[bidi=default]{babel}
\babelprovide[main,import]{english}

\babelfont{rm}[,RawFeature={fallback=mainfontfallback}]{CMU Serif}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

%\newlength\Oldarrayrulewidth
%\newlength\Oldtabcolsep

\usepackage{hyperref}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[nolongtablepatch]{lineno}
\linenumbers

\usepackage{fontspec}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

  \setmainfont[,RawFeature={fallback=mainfontfallback}]{CMU Serif}

\title{Modelling M/EEG data with Bayesian generalised additive
multilevel models}

\shorttitle{Bayesian M/EEG modelling}

\usepackage{etoolbox}

\authorsnames[{1},{2}]{Ladislas Nalborczyk,Paul Bürkner}

\authorsaffiliations{
{Aix Marseille Univ, CNRS, LPL},{TU Dortmund University, Department of
Statistics}}

\leftheader{Nalborczyk and Bürkner}

\abstract{Time-resolved electrophysiological measurements such as those
offered by magneto- or electro-encephalography (M/EEG) provide a unique
window onto neural activity underlying cognitive process and how they
unfold over time. Typically, we are interested in testing whether such
measures differ across conditions and/or groups. The conventional
approach consists in conducting mass-univariate statistics through
followed by some form of multiplicity correction (e.g., FDR, FWER) or
cluster-based inference. However, these cluster-based methods have an
important downside: they shift the focus of inference from the timepoint
to the cluster level, thus preventing any conclusion to be made about
the onset and offset of effects (e.g., differences across conditions).
Here, we introduce a novel \emph{model-based approch} for analysing
one-dimensional M/EEG timeseries such as ERPs or decoding timecourses
and their differences across conditions or group. This approach relies
on Bayesian generalised additive multilevel models, which output the
posterior probabilility of the effect being above 0 (or above chance) at
every timestep, while naturally taking into account the temporal
dependencies and between-subject variability present in such data.}

\keywords{EEG, MEG, generalised additive models, mixed-effects
models, multilevel models, Bayesian statistics, brms}

\authornote{\par{\addORCIDlink{Ladislas
Nalborczyk}{0000-0002-7419-9855}}\par{\addORCIDlink{Paul
Bürkner}{0000-0001-5765-8995}}

\par{   The authors have no conflicts of interest to disclose.    }
\par{Correspondence concerning this article should be addressed
to Ladislas Nalborczyk, Aix Marseille Univ, CNRS, LPL, 5 avenue
Pasteur, 13100
Aix-en-Provence, France, email: ladislas.nalborczyk@cnrs.fr}
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother

\urlstyle{same}

\usepackage{mathtools}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

%% Disable longtable counter
%% https://tex.stackexchange.com/a/248395/211326

\usepackage{etoolbox}

\makeatletter
\patchcmd{\LT@caption}
  {\bgroup}
  {\bgroup\global\LTpatch@captiontrue}
  {}{}
\patchcmd{\longtable}
  {\par}
  {\par\global\LTpatch@captionfalse}
  {}{}
\apptocmd{\endlongtable}
  {\ifLTpatch@caption\else\addtocounter{table}{-1}\fi}
  {}{}
\newif\ifLTpatch@caption
\makeatother

\begin{document}

\maketitle

\hypertarget{toc}{}
\tableofcontents
\newpage
\section[Introduction]{Modelling M/EEG data with Bayesian generalised
additive multilevel models}

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\resetlinenumber[1]

\section{Introduction}\label{introduction}

Here are some useful references to be discussed
(\citeproc{ref-combrisson_exceeding_2015}{Combrisson \& Jerbi, 2015};
\citeproc{ref-ehinger_unfold_2019}{Ehinger \& Dimigen, 2019};
\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021},
\citeproc{ref-frossard2022}{2022}; \citeproc{ref-gramfort2013}{Gramfort,
2013}; \citeproc{ref-hayasaka_validating_2003}{Hayasaka, 2003};
\citeproc{ref-luck_how_2017}{Luck \& Gaspelin, 2017};
\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al., 2019};
\citeproc{ref-pernet2015}{C. R. Pernet et al., 2015})\ldots{} See also
(\citeproc{ref-maris2011}{Maris, 2011})\ldots{} and
(\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}) (history of
cluster-based approaches and using a data split?)\ldots{} Cluster
failure (\citeproc{ref-eklund2016}{Eklund et al., 2016})\ldots{}

\subsection{Previous modelling work}\label{previous-modelling-work}

Recent example of GLM for EEG (\citeproc{ref-fischer2013}{Fischer \&
Ullsperger, 2013}; \citeproc{ref-wuxfcllhorst2025}{Wüllhorst et al.,
2025})\ldots{} ``Similar approaches have been successfully applied to
EEG time- (Rousselet et al., 2008) and frequency-domain (Cohen and
Cavanagh, 2011) data and allow the simultaneous investigation of
multiple independent variables while preserving the high temporal
resolution of the EEG.''\ldots{} See also (\citeproc{ref-hauk2006}{Hauk
et al., 2006}; \citeproc{ref-rousselet2008}{Rousselet et al.,
2008})\ldots{} Example of two-stage regression analysis (i.e.,
individual-level then group-level, \citeproc{ref-dunagan2024}{Dunagan et
al., 2024})\ldots{}

From Dimigen \& Ehinger (\citeproc{ref-dimigen2021}{2021}): ``Recently,
spline regression has been applied to ERPs (Hendrix, Baayen, \& Bolger,
2017; Kryuchkova,Tucker, Wurm, \& Baayen,2012; Tremblay \& Baayen, 2010;
Tremblay \& Newman, 2015)''\ldots{} GAMMs for EEG data
(\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
\citeproc{ref-meulman2015}{Meulman et al., 2015})\ldots{}

Disentangling overlapping processes
(\citeproc{ref-skukies_brain_2024}{Skukies et al., 2024};
\citeproc{ref-skukies_modelling_2021}{Skukies \& Ehinger, 2021})\ldots{}
Weighting single trials (\citeproc{ref-pernet2022}{C. Pernet,
2022})\ldots{} The LIMO toolbox
(https://onlinelibrary.wiley.com/doi/10.1155/2011/831409)\ldots{} Using
Bayes factors (\citeproc{ref-teichmann2022}{Teichmann, 2022})\ldots{}

\subsection{Generalised additive
models}\label{generalised-additive-models}

See for instance these tutorials
(\citeproc{ref-suxf3skuthy2017}{Sóskuthy, 2017};
\citeproc{ref-winter2016}{Winter \& Wieling, 2016}) or application to
phonetic data (\citeproc{ref-wieling2018}{Wieling, 2018}) or this
introduction (\citeproc{ref-baayen2020}{Baayen \& Linke, 2020}) or these
reference books (\citeproc{ref-hastie2017}{Hastie \& Tibshirani, 2017};
\citeproc{ref-wood2017}{Wood, 2017})\ldots{}

In generalised additive models (GAMs), the functional relationship
between predictors and response variable is decomposed into a sum of
low-dimensional non-parametric functions. A typical GAM has the
following form:

\[
\begin{aligned} 
y_{i} &\sim \mathrm{EF}\left(\mu_{i}, \phi\right)\\
g\left(\mu_i\right) &= A_{i} + \mathbf{X}_{i} \gamma + \sum_{j=1}^{J} f_{j}\left(x_{ij}\right)
\end{aligned}
\]

where \(y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)\) denotes that
the observations \(y_{i}\) are distributed as some member of the
exponential family of distributions (e.g., Gaussian, Gamma, Beta,
Poisson) with mean \(\mu_{i}\) and scale parameter \(\phi\);
\(g(\cdot)\) is the link function, \(A\) is an offset,
\(\mathbf{X}_{i}\) is the \(i\)th row of a parametric model matrix,
\(\gamma\) is a vector of parameters for the parametric terms, \(f_{j}\)
is a smooth function of covariate \(x_{j}\). The smooth functions
\(f_{j}\) are represented in the model via penalised splines basis
expansions of the covariates, that are a weighted sum of basis
functions:

\[
f_{j}\left(x_{i j}\right) = \sum_{k=1}^K \beta_{jk} b_{jk}\left(x_{ij}\right)
\]

where \(\beta_{jk}\) is the weight (coefficient) associated with the
\(k\)th basis function \(b_{jk}()\) evaluated at the covariate value
\(x_{ij}\) for the \(j\)th smooth function \(f_{j}\). Splines'
coefficients are penalised\ldots{}

\subsection{Objectives}\label{objectives}

Focusing on identifying onset and offset of effects (as assessed by ERP
differences or decoding performance)\ldots{} Assessing the performance
of a model-based approach (i.e., Bayesian GAMMs) to conventional methods
(multiplicity corrections or cluster-based permutation)\ldots{}

\section{Methods}\label{methods}

\subsection{M/EEG data simulation}\label{meeg-data-simulation}

Following the approach used by Sassenhagen \& Draschkow
(\citeproc{ref-sassenhagen2019}{2019}) and Rousselet
(\citeproc{ref-rousselet_using_2025}{2025}), we simulated EEG data
stemming from two conditions, one with noise only, and the other with
noise + signal. As in previous studies, the noise was generated by
superimposing 50 sinusoids at different frequencies, following an
EEG-like spectrum (see details and code in
\citeproc{ref-yeung2004}{Yeung et al., 2004}). As in Rousselet
(\citeproc{ref-rousselet_using_2025}{2025}), the signal was generated
from truncated Gaussian with an objective onset at 160 ms, a peak at 250
ms, and an offset at 342 ms. We simulated this signal for 250 timesteps
between 0 and 0.5s, akin to a 500 Hz sampling rate. We simulated such
data for a group of 20 participants with 50 trials per participant and
condition (Figure~\ref{fig-eeg}).

\begin{figure}[!htb]

\caption{\label{fig-eeg}Some ERPs in two conditions with 50 trials each,
for a group of 20 participants.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-eeg-1.pdf}

}

\end{figure}%

We computed the average of the ERP difference
(Figure~\ref{fig-erp})\ldots{}

\begin{figure}[!htb]

\caption{\label{fig-erp}Group-level average difference between
conditions (mean +/- standard error of the mean). The `true' onset and
offset are indicated by the vertical dashed lines.}

\centering{

\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-erp-1.pdf}

}

\end{figure}%

\subsection{Model fitting}\label{model-fitting}

We then fitted a Bayesian GAM using the \texttt{brms} package
(\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018};
\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}). We used the
default priors in \texttt{brms}, that is, weakly informative priors. We
ran eight Markov Chain Monte-Carlo (MCMC) to approximate the posterior
distribution, including each 5000 iterations and a warmup of 2000
iterations, yielding a total of \(8 \times (5000-2000) = 24000\)
posterior samples to use for inference. Posterior convergence was
assessed examining trace plots as well as the Gelman--Rubin statistic
\(\hat{R}\).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# averaging across participants}
\NormalTok{ppt\_df }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{group\_by}\NormalTok{(participant, condition, time) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{summarise}\NormalTok{(}\AttributeTok{eeg =} \FunctionTok{mean}\NormalTok{(eeg) ) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{ungroup}\NormalTok{()}

\CommentTok{\# defining a contrast for condition}
\FunctionTok{contrasts}\NormalTok{(ppt\_df}\SpecialCharTok{$}\NormalTok{condition) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{)}

\CommentTok{\# fitting the GAM}
\NormalTok{gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# cubic regression splines with k{-}1 basis functions}
\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
    \AttributeTok{data =}\NormalTok{ ppt\_df,}
    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/gam.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

\subsection{Posterior probability of difference above
0}\label{posterior-probability-of-difference-above-0}

We then plot the posterior predictions together with the posterior
estimate of the slope for \texttt{condition} at each timestep
(Figure~\ref{fig-plot-post-slope}).

\begin{figure}[!htb]

\caption{\label{fig-plot-post-slope}Posterior estimate of the ERP in
each condition (left) or directly for the difference of ERPs (right)
according to the GAM.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-plot-post-slope-1.pdf}

}

\end{figure}%

We then compute the posterior probability of the slope for
\texttt{condition} being above \(0+\epsilon\)
(Figure~\ref{fig-post-prob-test}), with \(\epsilon\) defined as 10\% of
the standard deviation of the raw EEG signal.

\begin{figure}[!htb]

\caption{\label{fig-post-prob-test}Posterior probability of the ERP
difference (slope) being above 0 according to the GAM.}

\centering{

\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-test-1.pdf}

}

\end{figure}%

We can also express this as the ratio of posterior probabilities (i.e.,
\(p/(1-p)\)) and visualise the timecourse of this ratio superimposed
with the conventional thresholds on evidence ratios
(Figure~\ref{fig-post-prob-ratio}).

\begin{figure}[!htb]

\caption{\label{fig-post-prob-ratio}Ratio of posterior probability
according to the GAM. Timesteps above threshold (10) are highlighted in
green.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.pdf}

}

\end{figure}%

\newpage

\subsection{Multilevel modelling using ERP summary
statistics}\label{multilevel-modelling-using-erp-summary-statistics}

Next we fit a hierarchical/multilevel GAM using summary statistics of
ERPs (mean and SD) at the participant level (similar to what is done in
meta-analysis).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# averaging across participants}
\NormalTok{summary\_df }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{summarise}\NormalTok{(}
        \AttributeTok{eeg\_mean =} \FunctionTok{mean}\NormalTok{(eeg),}
        \AttributeTok{eeg\_sd =} \FunctionTok{sd}\NormalTok{(eeg),}
        \AttributeTok{.by =} \FunctionTok{c}\NormalTok{(participant, condition, time)}
\NormalTok{        )}

\CommentTok{\# defining a contrast for condition}
\FunctionTok{contrasts}\NormalTok{(summary\_df}\SpecialCharTok{$}\NormalTok{condition) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{)}

\CommentTok{\# fitting the GAM}
\NormalTok{meta\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# using by{-}participant SD of ERPs across trials}
\NormalTok{    eeg\_mean }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
\NormalTok{        (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ participant),}
    \AttributeTok{data =}\NormalTok{ summary\_df,}
    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/meta\_gam.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

\subsection{Error properties of the proposed
approach}\label{error-properties-of-the-proposed-approach}

We then computed the difference between the true and estimated
onset/offset of the ERP difference
(\(\text{error}:=|\hat{\theta}-\theta|\)), according to various
\texttt{eps} and \texttt{threshold} values. Remember that the signal is
generated from a truncated Gaussian with an objective onset at 160 ms, a
maximum at 250 ms, and an offset at 342 ms. Figure~\ref{fig-onset-error}
shows that the hierarchical GAM can \emph{exactly} recover the true
onset and offset values, given some reasonable choice of \texttt{eps}
and \texttt{threshold} values (e.g., a threshold of 20).

\begin{verbatim}
   eps threshold estimated_onset estimated_offset error_onset error_offset
1 0.00        13            0.16             0.34           0        0.002
2 0.00        14            0.16             0.34           0        0.002
3 0.00        15            0.16             0.34           0        0.002
4 0.00        16            0.16             0.34           0        0.002
5 0.00        17            0.16             0.34           0        0.002
6 0.01        10            0.16             0.34           0        0.002
\end{verbatim}

\begin{figure}[!htb]

\caption{\label{fig-onset-error}Error function of onset (left) and
offset (right) estimation according to various eps and threshold values
(according to the hierarchical GAM). Minimum error values are indicated
by red crosses.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-error-1.pdf}

}

\end{figure}%

\newpage

\subsection{Comparing the identified onsets/offsets to other
approaches}\label{comparing-the-identified-onsetsoffsets-to-other-approaches}

We compared the ability of the GAMM to correctly estimate the onset and
offset of the ERP difference to widely used methods. First, we conducted
mass-univariate t-tests (thus treating each timestep independently) and
identified the onset and offset of the ERP difference as the first and
last values crossing an arbitrary significance threshold
(\(\alpha = 0.01\)). We then followed the same approach but after
applying different forms of multiplicity correction the the p-values. We
compared two methods that control the false discovery rate (FDR) (i.e.,
BH95, \citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995}; and
BY01, \citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001}), one
method that controls the familywise error rate (FWER) (i.e.,
Holm--Bonferroni method, https://www.jstor.org/stable/4615733), and two
cluster-based methods (permutation with a single threshold and TFCE,
\citeproc{ref-smith2009}{Smith \& Nichols, 2009}). The BH95, BY01, and
Holm corrections were applied to the p-values using the
\texttt{p.adjust()} function in \texttt{R}. The cluster-based inference
was implemented using a cluster-sum statistic of squared t-values, as
implemented in \texttt{MNE-Python}
(\citeproc{ref-gramfort2013}{Gramfort, 2013}), called via the \texttt{R}
package \texttt{reticulate} v 1.35.0 (\citeproc{ref-reticulate}{Ushey et
al., 2024}). We also compared these estimates to the onset and offset as
estimated using the binary segmentation algorithm, as implemented in the
\texttt{R} package \texttt{changepoint} v 2.2.4
(\citeproc{ref-changepoint}{Killick et al., 2022a}), and applied
directly to the squared t-values (as in
\citeproc{ref-rousselet_using_2025}{Rousselet, 2025}). For visualisation
and interpretability purposes, we converted p-values to s-values, which
can be interpreted as bits of surprising information, assuming a null
effect (\citeproc{ref-greenland2019}{Greenland, 2019})
(Figure~\ref{fig-corrections}).

\begin{figure}[!htb]

\caption{\label{fig-corrections}Timecourse of squared t-values and
s-values, with true onset (black dashed line) and onsets identified
using the raw (uncorrected) p-values or the corrected p-values (BH, BY,
Holm).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-corrections-1.pdf}

}

\end{figure}%

\newpage

\subsection{Simulation study}\label{simulation-study}

Onset/offset estimation methods were assessed using these summary
statistics of the Monte Carlo sampling distributions of onsets: median
absolute error (MAE) and variance, with 10.000 iterations (i.e., 10.000
simulated datasets)\ldots{}

\subsection{Application to actual MEG
data}\label{application-to-actual-meg-data}

Assessing the reliability of the proposed approach using some sort of
split-half reliability (e.g., \citeproc{ref-rosenblatt2018}{Rosenblatt
et al., 2018})? Using the MEG decoding results of Nalborczyk et al.~(in
preparation) in 33 participants:

\begin{itemize}
\tightlist
\item
  Create many (e.g., 10 or 20) half train/test splits of the data
\item
  For each fold, estimate the onset/offset on both splits using all
  methods
\item
  Then summarise the distribution of between-split difference across
  folds
\end{itemize}

This will allow checking that the proposed approach produces more
reliable onset/offset estimates.

\section{Results}\label{results}

\ldots{}

\subsection{Simulation results (bias and
variance)}\label{simulation-results-bias-and-variance}

Figure~\ref{fig-simulation-mae-variance} shows a summary of the
simulation results, revealing that the proposed approach (\texttt{brms})
has the lowest MAE and variance for both the onset and offset
estimates\ldots{}

\begin{figure}[!htb]

\caption{\label{fig-simulation-mae-variance}Median absolute error and
variance of onset and offset estimates for each method.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-simulation-mae-variance-1.pdf}

}

\end{figure}%

\begin{figure}[!htb]

\caption{\label{fig-simulation-distribution}Distributions of onset and
offset estimates for each method.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-simulation-distribution-1.pdf}

}

\end{figure}%

\subsection{Application to actual EEG data
(reliability)}\label{application-to-actual-eeg-data-reliability}

\ldots{}

\newpage

\section{Discussion}\label{discussion}

\ldots{}

\subsection{Summary of the proposed
approach}\label{summary-of-the-proposed-approach}

\ldots{}

\subsection{Increasing potential
usage}\label{increasing-potential-usage}

Prepare a wrapper \texttt{R} package and show how to call it in
\texttt{Python} and integrate it with \texttt{MNE-Python}
(\citeproc{ref-gramfort2013}{Gramfort, 2013}) pipelines\ldots{}

\subsection{Limitations and future
directions}\label{limitations-and-future-directions}

Can be applied to any 1D timeseries (e.g., pupillometry,
electromyography)\ldots{} Extending the approach to spatiotemporal data
(i.e., time + sensors)\ldots{}

The error properties depend on the threshold parameter, a value of 10 or
20 seems to be a reasonable default, but the optimal threshold parameter
can be adjusted using split-half reliability assessment\ldots{}

\subsection{Conclusions}\label{conclusions}

\ldots{}

\newpage

\section{Data and code availability}\label{data-and-code-availability}

The simulation results as well as the \texttt{R} code to reproduce the
simulations are available on GitHub:
\url{https://github.com/lnalborczyk/brms_meeg}.

\section{Packages}\label{packages}

We used R version 4.2.3 (\citeproc{ref-base}{R Core Team, 2023}) and the
following R packages: brms v. 2.22.0 (\citeproc{ref-brms2017}{Bürkner,
2017}, \citeproc{ref-brms2018}{2018}, \citeproc{ref-brms2021}{2021}),
changepoint v. 2.2.4 (\citeproc{ref-changepoint2022}{Killick et al.,
2022b}; \citeproc{ref-changepoint2014}{Killick \& Eckley, 2014}),
grateful v. 0.2.10 (\citeproc{ref-grateful}{Rodriguez-Sanchez \&
Jackson, 2023}), knitr v. 1.45 (\citeproc{ref-knitr2014}{Xie, 2014},
\citeproc{ref-knitr2015}{2015}, \citeproc{ref-knitr2023}{2023}),
MetBrewer v. 0.2.0 (\citeproc{ref-MetBrewer}{Mills, 2022}), pakret v.
0.2.2 (\citeproc{ref-pakret}{Gallou, 2024}), patchwork v. 1.2.0
(\citeproc{ref-patchwork}{T. L. Pedersen, 2024}), rmarkdown v. 2.29
(\citeproc{ref-rmarkdown2024}{Allaire et al., 2024};
\citeproc{ref-rmarkdown2018}{Xie et al., 2018},
\citeproc{ref-rmarkdown2020}{2020}), scales v. 1.3.0
(\citeproc{ref-scales}{Wickham et al., 2023}), scico v. 1.5.0
(\citeproc{ref-scico}{T. L. Pedersen \& Crameri, 2023}), tidybayes v.
3.0.6 (\citeproc{ref-tidybayes}{Kay, 2023}), tidyverse v. 2.0.0
(\citeproc{ref-tidyverse}{Wickham et al., 2019}).

\newpage

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-abugaber2023}
Abugaber, D., Finestrat, I., Luque, A., \& Morgan-Short, K. (2023).
Generalized additive mixed modeling of EEG supports dual-route accounts
of morphosyntax in suggesting no word frequency effects on processing of
regular grammatical forms. \emph{Journal of Neurolinguistics},
\emph{67}, 101137.
\url{https://doi.org/10.1016/j.jneuroling.2023.101137}

\bibitem[\citeproctext]{ref-rmarkdown2024}
Allaire, J., Xie, Y., Dervieux, C., McPherson, J., Luraschi, J., Ushey,
K., Atkins, A., Wickham, H., Cheng, J., Chang, W., \& Iannone, R.
(2024). \emph{{rmarkdown}: Dynamic documents for r}.
\url{https://github.com/rstudio/rmarkdown}

\bibitem[\citeproctext]{ref-baayen2020}
Baayen, R. H., \& Linke, M. (2020). \emph{Generalized Additive Mixed
Models} (pp. 563--591). Springer International Publishing.
\url{https://doi.org/10.1007/978-3-030-46216-1_23}

\bibitem[\citeproctext]{ref-benjamini1995}
Benjamini, Y., \& Hochberg, Y. (1995). Controlling the False Discovery
Rate: A Practical and Powerful Approach to Multiple Testing.
\emph{Journal of the Royal Statistical Society Series B: Statistical
Methodology}, \emph{57}(1), 289--300.
\url{https://doi.org/10.1111/j.2517-6161.1995.tb02031.x}

\bibitem[\citeproctext]{ref-benjamini2001}
Benjamini, Y., \& Yekutieli, D. (2001). The control of the false
discovery rate in multiple testing under dependency. \emph{The Annals of
Statistics}, \emph{29}(4). \url{https://doi.org/10.1214/aos/1013699998}

\bibitem[\citeproctext]{ref-brms2017}
Bürkner, P.-C. (2017). {brms}: An {R} package for {Bayesian} multilevel
models using {Stan}. \emph{Journal of Statistical Software},
\emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}

\bibitem[\citeproctext]{ref-brms2018}
Bürkner, P.-C. (2018). Advanced {Bayesian} multilevel modeling with the
{R} package {brms}. \emph{The R Journal}, \emph{10}(1), 395--411.
\url{https://doi.org/10.32614/RJ-2018-017}

\bibitem[\citeproctext]{ref-brms2021}
Bürkner, P.-C. (2021). Bayesian item response modeling in {R} with
{brms} and {Stan}. \emph{Journal of Statistical Software},
\emph{100}(5), 1--54. \url{https://doi.org/10.18637/jss.v100.i05}

\bibitem[\citeproctext]{ref-combrisson_exceeding_2015}
Combrisson, E., \& Jerbi, K. (2015). Exceeding chance level by chance:
{The} caveat of theoretical chance levels in brain signal classification
and statistical assessment of decoding accuracy. \emph{Journal of
Neuroscience Methods}, \emph{250}, 126--136.
\url{https://doi.org/10.1016/j.jneumeth.2015.01.010}

\bibitem[\citeproctext]{ref-dimigen2021}
Dimigen, O., \& Ehinger, B. V. (2021). Regression-based analysis of
combined EEG and eye-tracking data: Theory and applications.
\emph{Journal of Vision}, \emph{21}(1), 3.
\url{https://doi.org/10.1167/jov.21.1.3}

\bibitem[\citeproctext]{ref-dunagan2024}
Dunagan, D., Jordan, T., Hale, J. T., Pylkkänen, L., \& Chacón, D. A.
(2024). \emph{Evaluating the timecourses of morpho-orthographic,
lexical, and grammatical processing following rapid parallel visual
presentation: An EEG investigation in english}.
\url{http://dx.doi.org/10.1101/2024.04.10.588861}

\bibitem[\citeproctext]{ref-ehinger_unfold_2019}
Ehinger, B. V., \& Dimigen, O. (2019). Unfold: An integrated toolbox for
overlap correction, non-linear modeling, and regression-based {EEG}
analysis. \emph{PeerJ}, \emph{7}, e7838.
\url{https://doi.org/10.7717/peerj.7838}

\bibitem[\citeproctext]{ref-eklund2016}
Eklund, A., Nichols, T. E., \& Knutsson, H. (2016). Cluster failure: Why
fMRI inferences for spatial extent have inflated false-positive rates.
\emph{Proceedings of the National Academy of Sciences}, \emph{113}(28),
7900--7905. \url{https://doi.org/10.1073/pnas.1602413113}

\bibitem[\citeproctext]{ref-fischer2013}
Fischer, Adrian~G., \& Ullsperger, M. (2013). Real and Fictive Outcomes
Are Processed Differently but Converge on a Common Adaptive Mechanism.
\emph{Neuron}, \emph{79}(6), 1243--1255.
\url{https://doi.org/10.1016/j.neuron.2013.07.006}

\bibitem[\citeproctext]{ref-frossard2021}
Frossard, J., \& Renaud, O. (2021). Permutation Tests for Regression,
ANOVA, and Comparison of Signals: The {\textbf{permuco}} Package.
\emph{Journal of Statistical Software}, \emph{99}(15).
\url{https://doi.org/10.18637/jss.v099.i15}

\bibitem[\citeproctext]{ref-frossard2022}
Frossard, J., \& Renaud, O. (2022). The cluster depth tests: Toward
point-wise strong control of the family-wise error rate in massively
univariate tests with application to M/EEG. \emph{NeuroImage},
\emph{247}, 118824.
\url{https://doi.org/10.1016/j.neuroimage.2021.118824}

\bibitem[\citeproctext]{ref-pakret}
Gallou, A. (2024). \emph{{pakret}: Cite {``{R}''} packages on the fly in
{``{R Markdown}''} and {``{Quarto}''}}.
\url{https://CRAN.R-project.org/package=pakret}

\bibitem[\citeproctext]{ref-gramfort2013}
Gramfort, A. (2013). MEG and EEG data analysis with MNE-python.
\emph{Frontiers in Neuroscience}, \emph{7}.
\url{https://doi.org/10.3389/fnins.2013.00267}

\bibitem[\citeproctext]{ref-greenland2019}
Greenland, S. (2019). Valid {\emph{P}} -Values Behave Exactly as They
Should: Some Misleading Criticisms of {\emph{P}} -Values and Their
Resolution With {\emph{S}} -Values. \emph{The American Statistician},
\emph{73}(sup1), 106--114.
\url{https://doi.org/10.1080/00031305.2018.1529625}

\bibitem[\citeproctext]{ref-hastie2017}
Hastie, T. J., \& Tibshirani, R. J. (2017). \emph{Generalized Additive
Models}. Routledge. \url{https://doi.org/10.1201/9780203753781}

\bibitem[\citeproctext]{ref-hauk2006}
Hauk, O., Davis, M. H., Ford, M., Pulvermüller, F., \& Marslen-Wilson,
W. D. (2006). The time course of visual word recognition as revealed by
linear regression analysis of ERP data. \emph{NeuroImage}, \emph{30}(4),
1383--1400. \url{https://doi.org/10.1016/j.neuroimage.2005.11.048}

\bibitem[\citeproctext]{ref-hayasaka_validating_2003}
Hayasaka, S. (2003). Validating cluster size inference: Random field and
permutation methods. \emph{NeuroImage}, \emph{20}(4), 2343--2356.
\url{https://doi.org/10.1016/j.neuroimage.2003.08.003}

\bibitem[\citeproctext]{ref-tidybayes}
Kay, M. (2023). \emph{{tidybayes}: Tidy data and geoms for {Bayesian}
models}. \url{https://doi.org/10.5281/zenodo.1308151}

\bibitem[\citeproctext]{ref-changepoint2014}
Killick, R., \& Eckley, I. A. (2014). {changepoint}: An {R} package for
changepoint analysis. \emph{Journal of Statistical Software},
\emph{58}(3), 1--19.
\url{https://www.jstatsoft.org/article/view/v058i03}

\bibitem[\citeproctext]{ref-changepoint}
Killick, R., Haynes, K., \& Eckley, I. A. (2022a). \emph{{changepoint}:
An {R} package for changepoint analysis}.
\url{https://CRAN.R-project.org/package=changepoint}

\bibitem[\citeproctext]{ref-changepoint2022}
Killick, R., Haynes, K., \& Eckley, I. A. (2022b). \emph{{changepoint}:
An {R} package for changepoint analysis}.
\url{https://CRAN.R-project.org/package=changepoint}

\bibitem[\citeproctext]{ref-king2014}
King, J.-R., \& Dehaene, S. (2014). Characterizing the dynamics of
mental representations: the temporal generalization method. \emph{Trends
in Cognitive Sciences}, \emph{18}(4), 203--210.
\url{https://doi.org/10.1016/j.tics.2014.01.002}

\bibitem[\citeproctext]{ref-luck_how_2017}
Luck, S. J., \& Gaspelin, N. (2017). How to get statistically
significant effects in any {ERP} experiment (and why you shouldn't).
\emph{Psychophysiology}, \emph{54}(1), 146--157.
\url{https://doi.org/10.1111/psyp.12639}

\bibitem[\citeproctext]{ref-maris2011}
Maris, E. (2011). Statistical testing in electrophysiological studies.
\emph{Psychophysiology}, \emph{49}(4), 549--565.
\url{https://doi.org/10.1111/j.1469-8986.2011.01320.x}

\bibitem[\citeproctext]{ref-maris2007}
Maris, E., \& Oostenveld, R. (2007). Nonparametric statistical testing
of EEG- and MEG-data. \emph{Journal of Neuroscience Methods},
\emph{164}(1), 177--190.
\url{https://doi.org/10.1016/j.jneumeth.2007.03.024}

\bibitem[\citeproctext]{ref-meulman2015}
Meulman, N., Wieling, M., Sprenger, S. A., Stowe, L. A., \& Schmid, M.
S. (2015). Age Effects in L2 Grammar Processing as Revealed by ERPs and
How (Not) to Study Them. \emph{PLOS ONE}, \emph{10}(12), e0143328.
\url{https://doi.org/10.1371/journal.pone.0143328}

\bibitem[\citeproctext]{ref-MetBrewer}
Mills, B. R. (2022). \emph{{MetBrewer}: Color palettes inspired by works
at the metropolitan museum of art}.
\url{https://CRAN.R-project.org/package=MetBrewer}

\bibitem[\citeproctext]{ref-nalborczyk2019}
Nalborczyk, L., Batailler, C., Lœvenbruck, H., Vilain, A., \& Bürkner,
P.-C. (2019). An Introduction to Bayesian Multilevel Models Using brms:
A Case Study of Gender Effects on Vowel Variability in Standard
Indonesian. \emph{Journal of Speech, Language, and Hearing Research},
\emph{62}(5), 1225--1242.
\url{https://doi.org/10.1044/2018_jslhr-s-18-0006}

\bibitem[\citeproctext]{ref-pedersen_hierarchical_2019}
Pedersen, E. J., Miller, D. L., Simpson, G. L., \& Ross, N. (2019).
Hierarchical generalized additive models in ecology: An introduction
with mgcv. \emph{PeerJ}, \emph{7}, e6876.
\url{https://doi.org/10.7717/peerj.6876}

\bibitem[\citeproctext]{ref-patchwork}
Pedersen, T. L. (2024). \emph{{patchwork}: The composer of plots}.
\url{https://CRAN.R-project.org/package=patchwork}

\bibitem[\citeproctext]{ref-scico}
Pedersen, T. L., \& Crameri, F. (2023). \emph{{scico}: Colour palettes
based on the scientific colour-maps}.
\url{https://CRAN.R-project.org/package=scico}

\bibitem[\citeproctext]{ref-pernet2022}
Pernet, C. (2022). Electroencephalography robust statistical linear
modelling using a single weight per trial. \emph{Aperture Neuro},
\emph{2}, 1--22.
\url{https://doi.org/10.52294/apertureneuro.2022.2.seoo9435}

\bibitem[\citeproctext]{ref-pernet2015}
Pernet, C. R., Latinus, M., Nichols, T. E., \& Rousselet, G. A. (2015).
Cluster-based computational methods for mass univariate analyses of
event-related brain potentials/fields: A simulation study. \emph{Journal
of Neuroscience Methods}, \emph{250}, 85--93.
\url{https://doi.org/10.1016/j.jneumeth.2014.08.003}

\bibitem[\citeproctext]{ref-base}
R Core Team. (2023). \emph{{R}: A language and environment for
statistical computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\bibitem[\citeproctext]{ref-grateful}
Rodriguez-Sanchez, F., \& Jackson, C. P. (2023). \emph{{grateful}:
Facilitate citation of r packages}.
\url{https://pakillo.github.io/grateful/}

\bibitem[\citeproctext]{ref-rosenblatt2018}
Rosenblatt, J. D., Finos, L., Weeda, W. D., Solari, A., \& Goeman, J. J.
(2018). All-Resolutions Inference for brain imaging. \emph{NeuroImage},
\emph{181}, 786--796.
\url{https://doi.org/10.1016/j.neuroimage.2018.07.060}

\bibitem[\citeproctext]{ref-rousselet_using_2025}
Rousselet, G. A. (2025). Using cluster-based permutation tests to
estimate {MEG}/{EEG} onsets: {How} bad is it? \emph{European Journal of
Neuroscience}, \emph{61}(1), e16618.
\url{https://doi.org/10.1111/ejn.16618}

\bibitem[\citeproctext]{ref-rousselet2008}
Rousselet, G. A., Pernet, C. R., Bennett, P. J., \& Sekuler, A. B.
(2008). Parametric study of EEG sensitivity to phase noise during face
processing. \emph{BMC Neuroscience}, \emph{9}(1).
\url{https://doi.org/10.1186/1471-2202-9-98}

\bibitem[\citeproctext]{ref-sassenhagen2019}
Sassenhagen, J., \& Draschkow, D. (2019). Cluster{-}based permutation
tests of MEG/EEG data do not establish significance of effect latency or
location. \emph{Psychophysiology}, \emph{56}(6).
\url{https://doi.org/10.1111/psyp.13335}

\bibitem[\citeproctext]{ref-skukies_modelling_2021}
Skukies, R., \& Ehinger, B. (2021). Modelling event duration and overlap
during {EEG} analysis. \emph{Journal of Vision}, \emph{21}(9), 2037.
\url{https://doi.org/10.1167/jov.21.9.2037}

\bibitem[\citeproctext]{ref-skukies_brain_2024}
Skukies, R., Schepers, J., \& Ehinger, B. (2024, December 9).
\emph{Brain responses vary in duration - modeling strategies and
challenges}. \url{https://doi.org/10.1101/2024.12.05.626938}

\bibitem[\citeproctext]{ref-smith2009}
Smith, S., \& Nichols, T. (2009). Threshold-free cluster enhancement:
Addressing problems of smoothing, threshold dependence and localisation
in cluster inference. \emph{NeuroImage}, \emph{44}(1), 83--98.
\url{https://doi.org/10.1016/j.neuroimage.2008.03.061}

\bibitem[\citeproctext]{ref-suxf3skuthy2017}
Sóskuthy, M. (2017). \emph{Generalised additive mixed models for dynamic
analysis in linguistics: A practical introduction}.
\url{https://doi.org/10.48550/ARXIV.1703.05339}

\bibitem[\citeproctext]{ref-teichmann2022}
Teichmann, L. (2022). An empirically driven guide on using bayes factors
for m/EEG decoding. \emph{Aperture Neuro}, \emph{2}, 1--10.
\url{https://doi.org/10.52294/apertureneuro.2022.2.maoc6465}

\bibitem[\citeproctext]{ref-reticulate}
Ushey, K., Allaire, J., \& Tang, Y. (2024). \emph{Reticulate: Interface
to 'python'}. \url{https://CRAN.R-project.org/package=reticulate}

\bibitem[\citeproctext]{ref-tidyverse}
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}

\bibitem[\citeproctext]{ref-scales}
Wickham, H., Pedersen, T. L., \& Seidel, D. (2023). \emph{{scales}:
Scale functions for visualization}.
\url{https://CRAN.R-project.org/package=scales}

\bibitem[\citeproctext]{ref-wieling2018}
Wieling, M. (2018). Analyzing dynamic phonetic data using generalized
additive mixed modeling: A tutorial focusing on articulatory differences
between L1 and L2 speakers of English. \emph{Journal of Phonetics},
\emph{70}, 86--116. \url{https://doi.org/10.1016/j.wocn.2018.03.002}

\bibitem[\citeproctext]{ref-winter2016}
Winter, B., \& Wieling, M. (2016). How to analyze linguistic change
using mixed models, Growth Curve Analysis and Generalized Additive
Modeling. \emph{Journal of Language Evolution}, \emph{1}(1), 7--18.
\url{https://doi.org/10.1093/jole/lzv003}

\bibitem[\citeproctext]{ref-wood2003}
Wood, S. N. (2003). Thin Plate Regression Splines. \emph{Journal of the
Royal Statistical Society Series B: Statistical Methodology},
\emph{65}(1), 95--114. \url{https://doi.org/10.1111/1467-9868.00374}

\bibitem[\citeproctext]{ref-wood2017}
Wood, S. N. (2017). \emph{Generalized Additive Models}. Chapman;
Hall/CRC. \url{https://doi.org/10.1201/9781315370279}

\bibitem[\citeproctext]{ref-wuxfcllhorst2025}
Wüllhorst, V., Wüllhorst, R., Overmeyer, R., \& Endrass, T. (2025).
Comprehensive Analysis of Event{-}Related Potentials of Response
Inhibition: The Role of Negative Urgency and Compulsivity.
\emph{Psychophysiology}, \emph{62}(2).
\url{https://doi.org/10.1111/psyp.70000}

\bibitem[\citeproctext]{ref-knitr2014}
Xie, Y. (2014). {knitr}: A comprehensive tool for reproducible research
in {R}. In V. Stodden, F. Leisch, \& R. D. Peng (Eds.),
\emph{Implementing reproducible computational research}. Chapman;
Hall/CRC.

\bibitem[\citeproctext]{ref-knitr2015}
Xie, Y. (2015). \emph{Dynamic documents with {R} and knitr} (2nd ed.).
Chapman; Hall/CRC. \url{https://yihui.org/knitr/}

\bibitem[\citeproctext]{ref-knitr2023}
Xie, Y. (2023). \emph{{knitr}: A general-purpose package for dynamic
report generation in r}. \url{https://yihui.org/knitr/}

\bibitem[\citeproctext]{ref-rmarkdown2018}
Xie, Y., Allaire, J. J., \& Grolemund, G. (2018). \emph{R markdown: The
definitive guide}. Chapman; Hall/CRC.
\url{https://bookdown.org/yihui/rmarkdown}

\bibitem[\citeproctext]{ref-rmarkdown2020}
Xie, Y., Dervieux, C., \& Riederer, E. (2020). \emph{R markdown
cookbook}. Chapman; Hall/CRC.
\url{https://bookdown.org/yihui/rmarkdown-cookbook}

\bibitem[\citeproctext]{ref-yeung2004}
Yeung, N., Bogacz, R., Holroyd, C. B., \& Cohen, J. D. (2004). Detection
of synchronized oscillations in the electroencephalogram: An evaluation
of methods. \emph{Psychophysiology}, \emph{41}(6), 822--832.
\url{https://doi.org/10.1111/j.1469-8986.2004.00239.x}

\end{CSLReferences}

\newpage

\appendix

\section{Application to time-resolved decoding results (accuracy over
time)}\label{application-to-time-resolved-decoding-results-accuracy-over-time}

We conducted time-resolved multivariate pattern analysis (MVPA), also
known as decoding. As a result, we have a timecourse of decoding
accuracies (e.g., ROC AUC), bounded between 0 and 1, per participant
(Figure~\ref{fig-mne-decoding})\ldots{}

\begin{figure}[!htb]

\caption{\label{fig-mne-decoding}Exemplary average timecourse of binary
decoding accuracy (ROC AUC) for each participant. Group-level average
decoding accuracy is depicted as a grey background density in each
panel.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-mne-decoding-1.pdf}

}

\end{figure}%

Now, we want to \emph{test} whether the group-level average decoding
accuracy is above chance (i.e., 0.5) at each timestep. We use a similar
GAM/GP as previously, but we replace the \(\mathrm{Normal}\) likelihood
function by a \(\mathrm{Beta}\) one to account for the bounded nature of
AUC values (between 0 and 1).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting the GAM}
\NormalTok{decoding\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
    \AttributeTok{data =}\NormalTok{ decoding\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/decoding\_gam.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting the GP}
\NormalTok{decoding\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gp}\NormalTok{(time, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
    \AttributeTok{data =}\NormalTok{ decoding\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/decoding\_gp.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{Posterior predictions of the GAM (left) and GP (right) fitted
on decoding accuracy over time.}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/decoding-preds-1.pdf}

}

\end{figure}%

Next, we plot the posterior probability of decoding accuracy being above
chance level (plus some epsilon) (Figure~\ref{fig-decoding-post}).

\begin{figure}[!htb]

\caption{\label{fig-decoding-post}Posterior probability of decoding
accuracy being above chance level according to the GAM (left) or the GP
(right).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-post-1.pdf}

}

\end{figure}%

\begin{figure}[!htb]

\caption{\label{fig-decoding-ratio}Ratio of posterior probabilities of
decoding accuracy being above chance level according to the GAM (left)
or the GP (right).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-ratio-1.pdf}

}

\end{figure}%

\newpage

\section{Application to 2D time-resolved decoding results
(cross-temporal
generalisation)}\label{application-to-2d-time-resolved-decoding-results-cross-temporal-generalisation}

Assume we have M/EEG data and we have conducted cross-temporal
generalisation analyses (\citeproc{ref-king2014}{King \& Dehaene,
2014}). As a result, we have a 2D matrix where each element contains the
decoding accuracy (e.g., ROC AUC) of a classifier trained at timestep
\(\text{training}_{i}\) and tested at timestep \(\text{testing}_{j}\)
(Figure~\ref{fig-sim-timegen}).

\begin{figure}[!htb]

\caption{\label{fig-sim-timegen}Exemplary (simulated) group-level
average cross-temporal generalisation matrix of decoding performance
(ROC AUC).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-timegen-1.pdf}

}

\end{figure}%

Now, we want to test whether and when decoding performance is above
chance level (0.5 for a binary decoding task). These two models are
computationally heavier to fit (more observations and 2D smooth
functions)\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting a GAM with two temporal dimensions}
\NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# 2D thin{-}plate spline (tp)}
    \CommentTok{\# auc \textasciitilde{} s(train\_time, test\_time, bs = "tp", k = 10),}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
    \AttributeTok{data =}\NormalTok{ timegen\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"}
\NormalTok{    )}

\CommentTok{\# fitting a GP with two temporal dimensions}
\CommentTok{\# timegen\_gp \textless{}{-} brm(}
\CommentTok{\#     auc \textasciitilde{} gp(train\_time, test\_time, k = 20),}
\CommentTok{\#     data = timegen\_data,}
\CommentTok{\#     family = Beta(),}
\CommentTok{\#     control = list(adapt\_delta = 0.95),}
\CommentTok{\#     iter = 2000,}
\CommentTok{\#     chains = 4,}
\CommentTok{\#     cores = 4,}
\CommentTok{\#     file = "models/timegen\_gp.rds"}
\CommentTok{\#     )}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{Posterior probability of decoding accuracy being above chance
level (2D GAM).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-timegen-post-preds-1.pdf}

}

\end{figure}%

\newpage

\section{Mathematical formulation of the bivariate
GAM}\label{mathematical-formulation-of-the-bivariate-gam}

To model cross-temporal generalisation matrices of decoding performance
(ROC AUC), we extended the initial (decoding) GAM to take into account
the bivariate temporal distribution of AUC values, thus producing
naturally smoothed estimates (timecourses) of AUC values and posterior
probabilities. This model can be written as follows:

\[
\begin{aligned}
\text{AUC}_{i} &\sim \mathrm{Beta}(\mu_{i}, \phi)\\
g(\mu_{i}) &= f \left(\text{train}_{i}, \text{test}_{i} \right)\\
\end{aligned}
\]

where we assume that AUC values come from a \(\mathrm{Beta}\)
distribution with two parameters \(\mu\) and \(\phi\). We can think of
\(f \left(\text{train}_{i}, \text{test}_{i} \right)\) as a surface (a
smooth function of two variables) that we can model using a
2-dimensional splines. Let
\(\mathbf{s}_{i} = \left(\text{train}_{i}, \text{test}_{i} \right)\) be
some pair of training and testing samples, and let
\(\mathbf{k}_{m} = \left(\text{train}_{m}, \text{test}_{m} \right)\)
denote the \(m^{\text{th}}\) knot in the domain of \(\text{train}_{i}\)
and \(\text{test}_{i}\). We can then express the smooth function as:

\[
f \left(\text{train}_{i}, \text{test}_{i} \right) = \alpha + \sum_{m=1}^M \beta_{m} b_{m} \left(\tilde{s}_{i}, \tilde{k}_{m} \right)
\]

Note that \(b_{m}(,)\) is a basis function that maps
\(R \times R \rightarrow R\). A popular bivariate basis function uses
\emph{thin-plate splines} (\citeproc{ref-wood2003}{Wood, 2003}), which
extend to \(\mathbf{s}_{i} \in \mathbb{R}^{d}\) and \(\partial l_{g}\)
penalties. These splines are designed to interpolate and approximate
smooth surfaces over two dimensions (hence the ``bivariate'' term). For
\(d=2\) dimensions and \(l=2\) (smoothness penalty involving second
order derivative):

\[
f \left(\tilde{s}_{i} \right) = \alpha + \beta_{1} x_{i} + \beta_{2} z_{i} +\sum_{m=1}^{M} \beta_{2+m} b_m\left(\tilde{s}_i, \tilde{k}_m\right)
\]

using the the radial basis function given by:

\[
b_m\left(\tilde{s}_i, \tilde{k}_m\right)=\left\|\tilde{s}_i-\tilde{k}_m\right\|^2 \log \left\|\tilde{s}_i-\tilde{k}_m\right\|
\]

where \(\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|\) is the Euclidean
distance between the covariate \(\mathbf{s}_{i}\) and the knot location
\(\mathbf{k}_{m}\).

\newpage

\section{Threshold-free cluster
enhancement}\label{threshold-free-cluster-enhancement}

Cluster-based permutation approaches require defining a cluster-forming
threshold (e.g., a t- or f-value) as the initial step of the algorithm.
As different cluster-forming thresholds lead to clusters with different
spatial or temporal extent, this threshold modulates the sensitivity of
the subsequent permutation test. The threshold-free cluster enhancement
method (TFCE) was introduced by Smith \& Nichols
(\citeproc{ref-smith2009}{2009}) to overcome this arbitrary threshold.

In brief, the TFCE method works as follows. Instead of picking an
arbitrary cluster-forming threshold (e.g., \(t=2\)), we try all (or
many) possible thresholds in a given range and check whether a given
timestep/voxel belongs to a significant cluster under any of the set of
thresholds\ldots{} Then, instead of using cluster mass, we use a
weighted average between the cluster extend (\(e\), how broad is the
cluster, that is, how many connected samples it contains) and the
cluster height (\(h\), how high is the cluster, that is, how large is
the test statistic) according to the formula:

\[
\text{TFCE} = \int_{h} e(h)^{E} h^{H} \mathrm{d}h
\]

Where\ldots{} the parameters \(E\) and \(H\) are set a priori and
control the influence of the extend and height on the TFCE. Then,
p-value for timestep/voxel \(i\) is computed by comparing it TFCE with
the null distribution of TFCE values. For each permuted signal, we keep
the maximal value over the whole signal for the null distribution of the
TFCE\ldots. But see Sassenhagen \& Draschkow
(\citeproc{ref-sassenhagen2019}{2019})\ldots{}

\newpage

\section{\texorpdfstring{Using the \texttt{R} package and integration
with
\texttt{MNE-Python}}{Using the R package and integration with MNE-Python}}\label{using-the-r-package-and-integration-with-mne-python}

Explain how to use the \texttt{R} package and to integrate it with
\texttt{MNE} epochs\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# to{-}do adding some code here...}
\end{Highlighting}
\end{Shaded}

\end{document}